Properties of winning Iterated Prisoner's Dilemma strategies.

### 2020-03-29 paper/main.tex
finalise conclusion

\begin{abstract}
From the celebrated performance of Tit for Tat, to the
introduction of the zero-determinant strategies, to the use of sophisticated
structures such as neural networks, the literature has been exploring the
performance of Iterated Prisoner's Dilemma strategies for decades. The results of the literature,
however, have been relying on a small number of somewhat arbitrarily selected
strategies in a very small number of tournaments,
casting doubt on the generalizability of results from statistical perspective.
This manuscript evaluates \numberofstrategies strategies in 45686
tournaments, presents the top performing strategies, and analyzes their
salient features. The results imply that there is not a single
strategy that performs well in any Iterated Prisoner's Dilemma interaction.
There are several properties, however, that heavily influence the best performing
strategies, these are: be nice, be provocable, be a little envious, be clever
and adapt to the environment.
\end{abstract}

@@

\begin{itemize}
    \item \st{Do not be envious}
    \item Be "nice"; Do not be the first to defect
    \item Reciprocate both cooperation and defection; Be provocable
    \item \st{Do not be too clever}
    \item Adapt to the environment; Adjust to the mean cooperator
\end{itemize}

@@

This manuscript has explored the performance of \numberofstrategies strategies of the Iterated
Prisoner's Dilemma in a large number of computer tournaments. The results of
the analysis demonstrated that, although for specific tournament types such as
standard, probabilistic ending tournaments and noisy tournaments with \(p_n <0.1\), dominant types of strategies exist
there is not a single dominant strategy, from the collection on strategies considered in this work, in an IPD competition if the environments
vary. Moreover, a strategy with a theory of mind should aim to adapt its behaviour
based on the mean and median cooperators and should in general not be too
cooperative.

The \numberofstrategies strategies used in this manuscript have been mainly for
the literature, and they have been accessible due to an open source software
called the Axelrod-Python library. The software was used to generate a total of
\numberofalltournaments computer tournaments results with different number of
strategies and different participants each time. The data collection was
described in Section~\ref{section:data_collection}. In
Section~\ref{section:top_performances}, the tournaments results were used to
present the top performances. The data set contained results from four different
settings, and these were also studied individually. In standard tournaments
complex strategies trained using reinforcement learning ranked in the top spots.
Some of these strategies ranked again in the top spots in probabilistic
ending tournaments when a \(p_e\) of less 0.1 was considered and in noisy tournaments
when \(p_n\) was less than 0.1. In probabilistic
ending tournaments \(p_e\) was designed to vary between 0 and 1. It was demonstrated
that for values larger than 0.1, as stated in the Folk Theorem, defecting strategies
were winning the tournaments because there was a high likelihood of the game
ending in the next turn. In tournaments with noise the median ranks of the top
15 strategies had the highest values and the \(r\) distributions were bimodal.
The top rank strategies were performing both well and bad, and this indicates
that in noisy tournaments there are not strategies that can guarantee winning.
However, if the probability of noise was constrained at 0.1 then strategies
designed for noisy tournaments indeed performed well.
Overall, the top ranked strategies differed from one tournament type to
another and the mechanism behind the winning strategies were all different, with some
exceptions which include trained (Evolved) and Meta strategies.

Section~\ref{section:evaluation_of_performance} covered an analysis of
performance based on several features associated with a strategy and with the
environments in which it was competing. The results of this analysis showed that a
strategy's characteristics such as whether or not it's stochastic, and the information it
used regarding the game had no effect on the strategy's success. The most
important features have been those that compared the strategy's behaviour to its
environment. The cooperating ratio of the strategy compared to the mean and
median cooperator was highlighted as the most important feature in the analysis.
More specifically, if a strategy were to enter a tournament with a theory of
mind of its environment it would choose to be the median cooperator in standard
tournaments, to cooperate 10\% of the time the median cooperator did in probabilistic ending tournaments and 
60\% in noisy and noisy probabilistic
tournaments. Lastly, if a strategy was aware of the opponents but not of the
setting of the tournament, a strategy would be more likely to be successful if
it were to identify the median cooperator and cooperated 67\% of the times that
they did.

>>>>

\begin{abstract}
From the celebrated performance of Tit for Tat, to the
introduction of the zero-determinant strategies, to the use of sophisticated
structures such as neural networks, the literature has been exploring the
performance of Iterated Prisoner's Dilemma strategies for decades. The results of the literature,
however, have been relying on a small number of somewhat arbitrarily selected
strategies in a very small number of tournaments,
casting doubt on the generalizability of results from statistical perspective.
This manuscript evaluates \numberofstrategies strategies in 45686
tournaments, presents the top performing strategies, and analyzes their
salient features. The results imply that there is not a single
strategy that performs well in any Iterated Prisoner's Dilemma interaction.
There are several properties, however, that heavily influence the best performing
strategies, these are: be nice, be provocable and contrite, be a little envious, be clever
and adapt to the environment.
\end{abstract}

@@

\begin{itemize}
    \item \st{Do not be envious} Be a little bit envious
    \item Be "nice"; Do not be the first to defect
    \item Reciprocate both cooperation and defection; Be provocable and contrite
    \item \st{Do not be too clever} Be clever
    \item Adapt to the environment; Adjust to the mean cooperator
\end{itemize}

@@

This manuscript has explored the performance of \numberofstrategies strategies
of the Iterated Prisoner's Dilemma in a large number of computer tournaments.
The \numberofstrategies strategies used in this manuscript have been mainly for
the literature, and they have been accessible due to an open source software
called the Axelrod-Python library. The software was used to generate a total of
\numberofalltournaments computer tournaments results with different number of
strategies and different participants each time. The data collection was
described in Section~\ref{section:data_collection}.

In
Section~\ref{section:top_performances}, the tournaments results were used to
present the top performances. The data set contained results from four different
settings, and these were also studied individually. In standard tournaments
complex strategies trained using reinforcement learning ranked in the top spots.
Some of these strategies ranked again in the top spots in probabilistic
ending tournaments when a \(p_e\) of less 0.1 was considered and in noisy tournaments
when \(p_n\) was less than 0.1. In probabilistic
ending tournaments \(p_e\) was designed to vary between 0 and 1. It was demonstrated
that for values larger than 0.1, as stated in the Folk Theorem, defecting strategies
were winning the tournaments because there was a high likelihood of the game
ending in the next turn. In tournaments with noise the median ranks of the top
15 strategies had the highest values and the \(r\) distributions were bimodal.
The top rank strategies were performing both well and bad, and this indicates
that in noisy tournaments there are not strategies that can guarantee winning.
However, if the probability of noise was constrained at 0.1 then strategies
designed for noisy tournaments indeed performed well.

Section~\ref{section:evaluation_of_performance} covered an analysis of
performance based on several features associated with a strategy and with the
environments in which it was competing. The results of this analysis showed that a
strategy's characteristics such as whether or not it's stochastic, and the information it
used regarding the game had no effect on the strategy's success. The most
important features have been those that compared the strategy's behaviour to its
environment. The cooperating ratio of the strategy compared to the mean
cooperator was highlighted as the most important feature in the analysis.
Other significant features included the SSE of a strategy and it's cooperating
rate after being in either state \(CC\) and \(CD\).

So what is the best way of playing the game? and is there a single dominant
strategy in the IPD? There is not a single strategy, from the collection of the
195 strategies, that has managed to will all the tournaments it competed in.
Even if on average a strategy ranked highly in a specific environment it did not
guarantee its success over the different tournament types. However, he results
of sections~\ref{section:top_performances}
and~\ref{section:evaluation_of_performance} have demonstrated that there are
properties associated with the success of strategies. A few the properties that
have been identified by this manuscript's analysis contradict the properties of
Axelrod~\cite{Axelrod1981}. Namely, in Section~\ref{section:top_performances} it
was shown that trained strategies and strategies that decided their actions
based on pre designed strategies to maximise their utility dominated several
tournaments across tournament types. Hinting that successful IPD strategies are
clever. Most of the successful strategies highlighted
in~Section~\ref{section:top_performances} were strategies that did open with a
cooperation.

Furthermore, in Section~\ref{section:top_performances}
and~\ref{section:evaluation_of_performance} it was shown that envious strategies
performed well. Though these were not the most envious strategies in the
tournaments (ZDs were included) these strategies benefited by being a bit
envious. From Section~\ref{section:evaluation_of_performance} it was concluded
that there is a significant importance in adapting to the environment, and more
specifically in this work, to the mean cooperator. This section also
demonstrated that a strategy should reciprocate, as suggested by Axelrod, but in
some environments, such as standard and noisy, it should relax it's readiness to
do so.

Thus, the five properties successful strategies need to have in a IPD competition
are: be nice, be provocable and contrite, be a little envious, be clever and adapt to the
environment.
